---
title: "Modern Computational Statistical Methods"
subtitle: "STAT8178-STAT7178"
author: "Macquarie University S1 2022"
date: "Week 9"
output: beamer_presentation
theme: CambridgeUS
colortheme: "dove"
includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(moments) # calculate moments 
library(boot) # Bootstrapping
library(purrr) # rdunif: Generate random sample from a discrete uniform distribution
```



# R session

Source: `https://bookdown.org/egarpor/PM-UC3M/npreg-npdens.html`.
Example: we consider the `faithful` dataset, which contains the duration of the eruption and the waiting time between eruptions for the Old Faithful geyser in Yellowstone National Park (USA).

---

# R Session
\scriptsize
```{r, fig.height=3, fig.width=6}
# Duration of eruption
faithE <- faithful$eruptions

# Default histogram: automatically chooses bins and uses absolute frequencies
histo <- hist(faithE)
```

---

# R Session
\scriptsize
```{r, fig.height=3, fig.width=6}
# Bins and bin counts
histo$breaks # Bk's
histo$counts # vk's
# With relative frequencies
hist(faithE, probability = TRUE)
```

---

# R Session
\scriptsize
```{r, fig.height=3, fig.width=6}
# Choosing the breaks
t0 <- min(faithE)
h <- 0.25
Bk <- seq(t0, max(faithE), by = h)
hist(faithE, probability = TRUE, breaks = Bk)
rug(faithE) # The sample
```

---

# R Session

We focus first on exploring the dependence on $t_0$
  with the next example, as it serves for motivating the next density estimator.
  
\scriptsize
```{r, echo=FALSE, fig.height=3, fig.width=6}
# Uniform sample
set.seed(1234567)
u <- runif(n = 100)

# t0 = 0, h = 0.2
Bk1 <- seq(0, 1, by = 0.2)

# t0 = -0.1, h = 0.2
Bk2 <- seq(-0.1, 1.1, by = 0.2)

# Comparison
par(mfrow = 1:2)
hist(u, probability = TRUE, breaks = Bk1, ylim = c(0, 1.5),
     main = "t0 = 0, h = 0.2")
rug(u)
abline(h = 1, col = 2)
hist(u, probability = TRUE, breaks = Bk2, ylim = c(0, 1.5),
     main = "t0 = -0.1, h = 0.2")
rug(u)
abline(h = 1, col = 2)
```

---

# R Session

A popular collection of simulation scenarios was given by Marron and Wand (1992) and are conveniently available through the package `nor1mix`.
  
\scriptsize
```{r, echo=FALSE, fig.height=3, fig.width=6}
library(nor1mix)
# Available models
#?nor1mix::MarronWand

# Simulating -- specify density with MW object
samp <- nor1mix::rnorMix(n = 500, obj = nor1mix::MW.nm9)
hist(samp, freq = FALSE)

# Density evaluation
x <- seq(-4, 4, length.out = 400)
lines(x, nor1mix::dnorMix(x = x, obj = nor1mix::MW.nm9), col = 2)


```

---

# Session R


\scriptsize
```{r, echo=FALSE}
# Plot a MW object directly
# A normal with the same mean and variance is plotted in dashed lines
par(mfrow = c(2, 2))
plot(nor1mix::MW.nm5)
plot(nor1mix::MW.nm7)
plot(nor1mix::MW.nm10)
plot(nor1mix::MW.nm12)
lines(nor1mix::MW.nm1, col = 2:3) # Also possible
```

---

# Session R

 Generate `n` random numbers, distributed according to a normal mixture

\scriptsize
```{r, echo=FALSE}
x <- rnorMix(5000, MW.nm10)
# hist(x)# you don't see the claw
plot(density(x), ylim = c(0,0.6),
main = "Estim. and true 'MW.nm10' density")
lines(MW.nm10, col = "orange")
```

---

# Example 1

Example 9.1. Computational Statistics Handbook with MATLAB, 3rd Edition
Wendy L. and Angel R. Martinez.
\scriptsize
```{r, echo=FALSE,fig.height=3, fig.width=6}
n=1000
x<- rnorm(n,1)

# Histogram Estimator

histog<- hist(x, probability = T)
#Get the bin width.
h<- histog$breaks[2]-histog$breaks[1]
# Now return an estimate at a point xo.
xo = 0
# Find all of the bin limits less than xo.
bc<- histog$breaks
ind = which(bc < xo)
# xo should be between these two bin centers.
end<-length(ind)
b1 = bc[ind[end]]
b2 = bc[ind[end]+1]
fhat<- histog$density[end]
fhat
```

---

# Example 2

Example 9.2
Computational Statistics Handbook with MATLAB, 3rd Edition
 Wendy L. and Angel R. Martinez.
 
 \scriptsize
```{r, echo=FALSE,fig.height=3, fig.width=6}
geyser <- read.csv("geyser.csv")
x= geyser$geyser
n = length(x)
# Use Normal Reference Rule for bin width.
h = 3.5*sd(x)*n^(-1/3)
# Get the bin mesh. Put some extra space on the ends.
t0 = min(x) - 1
tm = max(x) + 1
rng = tm - t0
nbin = ceiling(rng/h) #returns the smallest integers not less than x
bins = seq(t0, nbin*h + t0, by =h)
hist(x, probability = TRUE, breaks = bins, main = "Old Faithful - Waiting Time Between Eruptions", xlab = "Waiting Times (minutes)", ylab = "Probability")
rug(x) # The sample
```

---

# Reference:
\begin{itemize}

\item Martinez, W. L. and Martinez, A. R. (2016). Computational Statistics Handbook with MATLAB (third edition). CRC Press.

\end{itemize}
