---
title: "SGTA Week 8"
author: "Hassan Doosti"
date: "20/04/2021"
output:
  pdf_document: default
  html_document: default
---

## Question 1


```{r}
options(scipen=999) #prevent scientific notation
# Load up the data.
mcdata = read.csv("mcdata.csv")
x= mcdata$mcdata
  n = length(x) #sample size
mu <- 454 #Population mean
sigma = 7.8 #Population Standard Devision
```


```{r}
sigxbar <- sigma/sqrt(n)#Standard Error or Variance of Sample Mean
alpha<-0.05 #Size of Test
# Try the t-test once
Tobs = (mean(x)-mu)/sigxbar 
Tobs

```


```{r}
# Apply Monte Carlo Simulation
M=1000
 
# Critical value Result
T.Statistic=array(dim=M)

for(i in 1:M){
  
# Decide on a pseudo-population that reflects the characteristics
# of the true population under the null hypothesis.  

# Obtain a random sample of size n from the pseudo-population.  
x=rnorm(n,mu,sigma)

# Calculate the value of the test statistic using the random
# sample in step above and record it
T.Statistic[i]=(mean(x)-mu)/sigxbar
  
}
#T.Statistic
# critiacal values
cv1 = quantile(T.Statistic,alpha/2)
cv2 = quantile(T.Statistic,1-alpha/2)
cv1
cv2
```

```{r}
# Test Statistic
test.statistic <- Tobs
test.statistic
```

#Result: Estimated critical values are -1.885069 and 1.912626. Since the observed value of the test statistic −2.56 is less than the lower critical value we reject H0.

P-Values


```{r}
ind1 <- which(T.Statistic<= -abs(Tobs))
ind2 = which(T.Statistic >= abs(Tobs))
pvalhat = length(ind1)/M +length(ind2)/M
pvalhat
```
# Result: Estimated p-value is less than the significant level so we need to reject H0. 


## Question 2. Part 1


```{R}

xbar= 47.2
sigma=15
mu=45
n = 100
M<-1000

#Get the critical value, using z as test statistic.
cv <- qnorm(0.95, 0, 1) # True Critical Region

sigxbar <- sigma/sqrt(n)#Population sigma is known.
alpha<-0.05

#Get some alternative hypotheses:
mualt <- 46:55 
j<-length(mualt)
Tm=array(dim=M)

betahat=array(dim=j)

for(j in 1:length(mualt)){
  
Im =0

#Get the true mean.
 mu = mualt[j]
 
for(i in 1:M)  {
  
#Generate a sample from H_1.
 # xs = sigma*randn(1,n) + mu

xs=rnorm(n,mu,sigma) 
Tm[i]=(mean(xs)-45)/sigxbar
# Tm[i]=t.test(x, mu=mu)$statistic  

if(Tm[i]<cv) {
 # Then did not reject H_0.
  Im = Im +1

               }

 betahat[j]=Im/M 
             } 
                       }


plot(mualt,betahat)
```
# Result: As the mu is going up the probability of Type II Error is going down.




## Question 2. Part 2


Sample size is 50
```{R}
xbar= 47.2
sigma=15
mu=45
n = 50
M<-1000

#Get the critical value, using z as test statistic.
cv <- qnorm(0.95, 0, 1) # True Critical Region

sigxbar <- sigma/sqrt(n)#Population sigma is known.
alpha<-0.05

#Get some alternative hypotheses:
mualt <- 40:55 
j<-length(mualt)
Tm=array(dim=M)

betahat=array(dim=j)

for(j in 1:length(mualt)){
  
Im =0

#Get the true mean.
 mu = mualt[j]
 
for(i in 1:M)  {
  
#Generate a sample from H_1.
 # xs = sigma*randn(1,n) + mu

xs=rnorm(n,mu,sigma) 
Tm[i]=(mean(xs)-45)/sigxbar
# Tm[i]=t.test(x, mu=mu)$statistic  

if(Tm[i]<cv) {
 # Then did not reject H_0.
  Im = Im +1

               }

 betahat[j]=Im/M 
             } 
                       }
#Get the estimated power
 powhatn50   = 1-betahat 


plot(mualt,powhatn50)
```


Sample size is 100


```{R}
xbar= 47.2
sigma=15
mu=45
n = 100
M<-1000

#Get the critical value, using z as test statistic.
cv <- qnorm(0.95, 0, 1) # True Critical Region

sigxbar <- sigma/sqrt(n)#Population sigma is known.
alpha<-0.05

#Get some alternative hypotheses:
mualt <- 40:55 
j<-length(mualt)
Tm=array(dim=M)

betahat=array(dim=j)

for(j in 1:length(mualt)){
  
Im =0

#Get the true mean.
 mu = mualt[j]
 
for(i in 1:M)  {
  
#Generate a sample from H_1.
 # xs = sigma*randn(1,n) + mu

xs=rnorm(n,mu,sigma) 
Tm[i]=(mean(xs)-45)/sigxbar
# Tm[i]=t.test(x, mu=mu)$statistic  

if(Tm[i]<cv) {
 # Then did not reject H_0.
  Im = Im +1

               }

 betahat[j]=Im/M 
             } 
                       }
#Get the estimated power
 powhatn100   = 1-betahat 


plot(mualt,powhatn100)
```
Sample size is 200.
```{R}
xbar= 47.2
sigma=15
mu=45
n = 200
M<-1000

#Get the critical value, using z as test statistic.
cv <- qnorm(0.95, 0, 1) # True Critical Region

sigxbar <- sigma/sqrt(n)#Population sigma is known.
alpha<-0.05

#Get some alternative hypotheses:
mualt <- 40:55 
j<-length(mualt)
Tm=array(dim=M)

betahat=array(dim=j)

for(j in 1:length(mualt)){
  
Im =0

#Get the true mean.
 mu = mualt[j]
 
for(i in 1:M)  {
  
#Generate a sample from H_1.
 # xs = sigma*randn(1,n) + mu

xs=rnorm(n,mu,sigma) 
Tm[i]=(mean(xs)-45)/sigxbar
# Tm[i]=t.test(x, mu=mu)$statistic  

if(Tm[i]<cv) {
 # Then did not reject H_0.
  Im = Im +1

               }

 betahat[j]=Im/M 
             } 
                       }
#Get the estimated power
 powhatn200   = 1-betahat 


plot(mualt,powhatn200)
```


# Result: As sample size increases this variance shrinks to zero. The smaller our variance the more powerful we expect our test to be at avoiding type II errors (failing to reject H0).





